---
title: "Exercise using R with ggplot2 (hotel-bookings)"
author: "Lourdes Atilano"
format: 
   html:
     toc: true
     code-fold: show
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 5
)
```

## Practicing with ggplot2

In this activity, you will review a scenario, and use ggplot2 to quickly create data visualizations that allow you to explore your data and gain new insights. You will learn more about basic ggplot2 syntax and data visualization in R.

## The Scenario

In this scenario, you are a Data Analyst Jr. working for a hotel booking company. You have cleaned and manipulated your data, and gotten some initial insights you would like to share.

### Step 1: Import the data

```{r}
hotel_bookings <- read.csv("exports/tablas/hotel_bookings.csv")
```

### Step 2: Look at a sample of your data

```{r}
head(hotel_bookings)
```

Table with the info of the first six rows of the dataset

```{r}
 colnames(hotel_bookings)
```

The information about the names of all the columns of the dataset

### Step 3: Install and load the 'ggplot2' package

```{r}
 install.packages("ggplot2", repos = "http://cran.us.r-project.org")
 #install.packages('tidyverse', repos = "http://cran.us.r-project.org")
 library(ggplot2)
```

### Step 4: Graphics

#### Scatterplots
##### Hypothesis: Market Segment by children

Plot about the hypothesis if people with children have to book in advance or not

```{r}
 ggplot(data = hotel_bookings) +
  geom_point(mapping = aes(x = lead_time, y = children)) + 
  labs(title = "Hypothesis that people with children have to book in advance")
```

On the x-axis, the plot shows how far in advance a booking is made, with the bookings furthest to the right happening the most in advance. On the y-axis it shows how many children there are in a party.

#### Bar chats
##### Change from scatterplot to bars chat

Plot about which group of guests books the most weekend nights to target that group in a new marketing campaign. The hypothesis is whether guests without children book the most weekend nights.

```{r}
 ggplot(data = hotel_bookings) +
  geom_point(mapping = aes(x = stays_in_weekend_nights, y = children)) + 
  labs(title = "Hypothesis of the group of guests book the most weekend nights")
```

For me, it was easier to interpret this graphic by changing the type to a chart bar, so I created a variation with this result

```{r}
 ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(y = children)) +
  labs(title="Market Segment by children")
```
The same graphic but with vertical bars

```{r}
 ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = children)) +
  labs(title="Market Segment by children")
```

This graphic shows that it is indeed true: the segment of people without children books significantly more on weekends than those with families.

##### Hypothesis about promotions

Your stakeholder is interested in developing promotions based on different booking distributions, but first they need to know how many of the transactions are occurring for each different distribution type.

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel))
```

##### Hypothesis the Market Segment

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill=market_segment))
```

##### Hypothesis about the deposit types

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill=deposit_type))
```

#### Facets

##### This facet create a different chart for each deposit type:

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~deposit_type) +
  theme(axis.text.x = element_text(angle = 45))
```

##### This facet create a different chart for each market segment

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_wrap(~market_segment) +
  theme(axis.text.x = element_text(angle = 45))
```

The stakeholder wants to run a family-friendly promotion targeting key market segments. She wants to know which market segments generate the largest number of bookings, and where these bookings are made (city hotels or resort hotels). 

##### This graphic create a bar chart showing each hotel type and market segment

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = hotel, fill = market_segment))
```

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = hotel, fill=hotel)) +
  facet_wrap(~market_segment)
```

#### Filtering

```{r}
  library(tidyverse)
```

##### Hypothesis about a promotion to families that make online bookings for city hotels

```{r}
onlineta_city_hotels <- filter(hotel_bookings, 
                           (hotel=="City Hotel" & 
                             hotel_bookings$market_segment=="Online TA"))
```

The `filter()` function to create a data set that only includes the data you want. You can use the '&' character to demonstrate that you want two different conditions to be true. Also, you can use the '$' character to specify which column in the data frame 'hotel_bookings' you are referencing (for example, 'market_segment'). 

```{r}
View(onlineta_city_hotels)
```

Another way to do this, is using the pipe operator

```{r}
onlineta_city_hotels_v2 <- hotel_bookings %>%
  filter(hotel=="City Hotel") %>%
  filter(market_segment=="Online TA")
```

Notice how in the code chunk above, the %>% symbol is used to note the logical steps of this code. First, it starts with the name of the data frame, `onlineta_city_hotels_v2`, AND THEN it tells `R` to start with the original data frame `hotel_bookings`. Then it tells it to filter on the 'hotel' column; finally, it tells it to filter on the 'market_segment' column.

```{r}
ggplot(data = onlineta_city_hotels_v2) +
  geom_point(mapping = aes(x = lead_time, y = children))
```

This scatterplot shows data for online bookings for city hotels. The plot reveals that bookings with children tend to have a shorter lead time, and bookings with 3 children have a significantly shorter lead time (<200 days). So, promotions targeting families can be made closer to the valid booking dates.
